Evaluating infectious disease forecasts with allocation scoring rules

### 2023-09-02 alloscore_manuscript/alloscore-application.Rnw
rewording of step 3 and removal of lambda^star

\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^F$, is the allocation that minimizes the expected loss:
\begin{align}
    x^F &= \underset{0 \leq x}{\mathrm{arg min}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations. It can be shown that the Bayes act has $x^F_i = F_i^{-1}(1 - \lambda^{\star}/L)$, where $\lambda^{\star}$ depends on the problem parameters $K$ and $L$ as well as the forecast distributions and is chosen so as to satisfy the equation
\begin{align}
\sum_{i=1}^{N}F_i^{-1}(1 - \lambda^{\star}/L) = K.
\end{align}
This partial solution to the allocation problem seems to have first appeared in \cite{hadleywhitin1963}; see the supplemental materials for a derivation.

One interpretation of this result is that the Bayes act sets the allocation in each location $i$ to a quantile of the forecast distribution $F_i$ for that location. The quantile is at a probability level $\alpha = 1 - \lambda^{\star}/L$ that is the same for all locations, and is chosen such that the constraint is satisfied. According to this interpretation, we divide the available resources among the locations in such a way that the forecasted probability that need will exceed the available resources is the same in all locations.
An alternative interpretation comes from noting that for each location $i$, $\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^F} = - \lambda^{\star}$ (see the supplement for a proof).
In words, at the allocation given by the Bayes act, the rate of change of the expected score as a function of the amount allocated to location $i$ is given by $-\lambda^{\star}$.
This derivative is the same for all locations, so the optimal allocation divides the available resources across all locations in such a way that according to $F$, the expected benefit of 1 additional unit of resources is the same in all locations.

\paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as
\begin{equation}
S_A(F, y; K) = s_A(x^F, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(1 - \lambda^{\star}/L))
\end{equation}
This score measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.

@@

The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\alpha = 1 - \lambda^{\star}/L$. This is appropriate if the resource constraint $K$ is a known constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$

>>>>

\paragraph{Step 2: Given a probabilistic forecast $F$, identify the Bayes act.} The Bayes act associated with the forecast, $x^{F,K}$, is the allocation that minimizes the expected loss:
\begin{align}
    x^{F,K} &= \underset{0 \leq x}{\mathrm{arg min}}\,\, \mathbb{E}_{F} [s_A(x, Y)] \text{ subject to }
     \, \sum_{i=1}^N x_i = K, \label{AP}
\end{align}
where $\mathbb{E}_{F} [s_A(x, Y)] = \sum_{i=1}^{N} L \mathbb{E}_{F_i}[\max(0, Y_i - x_i)]$ sums the expected loss due to unmet need across all locations.

In the supplement we show that for a typical forecast $F$ and constraint level $K$, the components $x_i^{F,K}$ of the Bayes act are quantiles 
$q_{i,\tau} = F_i^{-1}(\tau)$ of the marginal forecasts $F_i$ for a probability level $\tau = \tau(K,F)$ satisfying the equation
\begin{align}
\sum_{i=1}^{N}q_{i,\tau} = K.
\end{align}
We note that $\tau(K,F)$ depends not only on the constraint level $K$ defining the decision problem but also on the forecast $F$. This is quite
different from the situation for quantile scores and the WIS, where the probability levels used to calculate scores are fixed.

The fact that the vector $(q_{1,\tau},\ldots,q_{N,\tau})$ of quantiles solves the allocation problem \eqref{AP} has the 
interpretation that 
in order to allocate optimally (according to $F$) we must divide resoruces among the locations so that the forecasted probability of need exceeding local allocation is the same everywhere. A related interpretation is offered by our model's expression as a partial derivative (derived in the supplement) for the reduction in expected score (i.e., increase in expected utility) in location $i$ that results from a unit increase in the resource:
\begin{align}
\frac{\partial}{\partial x_i} \mathbb{E}_{F} [s_A(x, Y)] \vert_{x=x^{F,K}} = L \cdot (1-\tau).
\end{align}
This equation tells us that when our allocations are set to the Bayes act $x^{F,K}$, shifting a small amount $\Delta x$ of the resource from one location to another would yield no change in expected utility since the gain $\Delta x \cdot L \cdot (1-\tau)$ in the receiving location would be canceled out by the loss $\Delta x \cdot L \cdot (\tau-1)$ in the location giving up the resource. 
This partial solution to the allocation problem is well-known in inventory management and is often attributed to \cite{hadleywhitin1963}. There, products and order quantities play the roles we have given to locations and resource allocation levels.

\paragraph{Step 3: Define the scoring rule.} We can now define a proper scoring rule for the probabilistic forecast $F$ as
\begin{equation}
S_A(F, y; K) = s_A(x^{F,K}, y) = \sum_{i=1}^n L \max(0, y_i - F_i^{-1}(\tau(K,L,F)))
\end{equation}
This score measures the total unmet need across all locations that results from using the Bayes allocation associated with the forecast $F$ when the actual level of need in each location is observed to be $y_i$.

@@

The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau$. This is appropriate if the resource constraint $K$ is a known constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
